---
title: "Model Report"
author: "Sam Abbott"
date: "`r Sys.Date()`"
output: html_document
params: 
  model: ""
  model_dir: "results/model-run-2018-09-07_12:11:07"
  nthreads: 1
vignette: >
  %\VignetteIndexEntry{Model Report}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}

---

```{r setup, include=FALSE}
if (!exists("model")) {
  model <- ""
}

if (!exists("nthread")) {
  nthreads <- future::availableCores()[[1]]
}

if (!exists("model_dir")) {
  model_dir <- "results/evaluated-scenarios-2018-10-04_14:21:44-calibration-run/baseline-2018-10-04_14:21:45"
}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, 
                      message = TRUE, cache = FALSE,
                      fig.path = file.path(model_dir, "plots"),
                      cache.path = file.path(model_dir, "report-cache"),
                      fig.align = 'center',
                      fig.length = 8, fig.width = 8)
```

# Set-up

- Load packages required for analysis

```{r, packages, message = FALSE}
library(tidyverse)
library(rbi)
library(rbi.helpers)
library(prettypublisher)
library(ModelTBBCGEngland)
```


- Load the fitted model and prior draws.

```{r load-model}
if (class(model) %in% "character"){
  model <- ModelTBBCGEngland::read_libbi(file.path(model_dir, "libbi", "posterior"))
}

model
```

# Scenarios

- Set up settings for prediction

```{r}
run_time <- 2040 - 1931
 
scenario_prediction <-partial(sample,
                      target = "prediction",
                      with = "transform-obs-to-state",
                      end_time = run_time, 
                      noutputs = run_time,
                      debug = FALSE,
                      nthreads = nthreads)
```
- Predict future for baseline scenario (school age vaccination).

```{r baseline}
baseline_scenario <- scenario_prediction(model)
```

- Predict future for no vaccination scenario

```{r no-vac}
no_vac_model <- model
no_vac_model$model <- no_vac_model$model %>% 
  fix(vac_scheme = 2)

no_vac_scenario <- scenario_prediction(no_vac_model)
```


- Predict future for neonatal vaccination

```{r neonatal}
neonatal_model <- model
neonatal_model$model <- neonatal_model$model %>% 
  fix(vac_scheme = 1)

neonatal_scenario <- scenario_prediction(neonatal_model)
```

- List all models and assign names

```{r}
models <- list(baseline_scenario, neonatal_scenario, no_vac_scenario)
names(models) <- c("School-age BCG", "Neonatal BCG", "No BCG")
```

# Generate summarised output

- Plot an overview of states 

```{r overview-scenarios, fig.height = 16, fig.width = 16}
plot_state(models, summarise = TRUE)
```

- Plot an overview of states zoomed into 2000 onwards

```{r overview-scenarios-2000, fig.height = 16, fig.width = 16}
plot_state(models, summarise = TRUE, start_time = 69)
```

- Plot fitted observed states.

```{r overview-scenarios-2000-obs}
plot_state(models, states = c("YearlyHistPInc", "YearlyInc"), start_time = 69)
```

- Table fitted states

```{r, eval = FALSE}
pred_obs <- plot_state(models, states = c("YearlyHistPInc", "YearlyInc", "YearlyAgeInc"), summarise = FALSE, start_time = 74, plot_data = FALSE)

obs <- ModelTBBCGEngland::setup_model_obs() %>% 
  bind_rows(.id = "state") %>% 
  mutate(time = time + 1931)

pred_obs <- pred_obs %>% 
  dplyr::filter(Average == "median") %>% 
  mutate(pred_value = pretty_ci(Count, lll, hhh, digits = 0)) %>% 
  left_join(obs, by = c("state" = "state", "time" = "time", "age" = "age")) %>% 
  select(Observation = state, Age = age, Year = time,  `Observed Incidence` = value, `Predicted Incidence` = pred_value) %>% 
  mutate(Year = Year + 1931)


sum_obs_table <- pred_obs %>% 
  dplyr::filter(Observation %in% c("YearlyInc", "YearlyHistPInc")) %>% 
  drop_na(`Observed Incidence`) %>% 
  select(-Age) %>% 
  mutate(Observation = case_when(Observation == "YearlyInc" ~ "UK born TB cases",
                                 Observation == "YearlyHistPInc" ~ "ALL Pulmonary TB cases"))


saveRDS(sum_obs_table, file.path(model_dir, "data", "sum_obs_table.rds"))

age_cases_table <- pred_obs %>% 
  dplyr::filter(Observation %in% c("YearlyAgeInc")) %>% 
  drop_na(`Observed Incidence`) %>% 
  select(-Observation) %>% 
  group_by(Year) %>% 
  mutate(Age = c(paste0(seq(0, 65, 5), "-", seq(4, 69, 5)), "70-89")) %>% 
  ungroup


saveRDS(age_cases_table, file.path(model_dir, "data", "age_cases_table.rds"))
```

```{r, eval = FALSE}
knitr::kable(sum_obs_table)
```

```{r, eval = FALSE}
knitr::kable(age_cases_table)
```

- Plot concentrated age distributed cases

```{r overview-scenarios-con-age}
plot_state(model, states = c( "YearlyChildInc", "YearlyAdultInc", "YearlyOlderAdultInc"), start_time = 69) + theme(legend.position = "none")
```

- Plot age distributed cases.

```{r overview-scenarios-age, fig.height = 16, fig.width = 16}
plot_state(model, states = c("YearlyAgeInc"), start_time = 69) + theme(legend.position = "none")
```
