---
title: "Model Report"
author: "Sam Abbott"
date: "`r Sys.Date()`"
output: html_document
params: 
  model: ""
  model_dir: "results/evaluated-scenarios-2018-09-21_14:09:50/baseline-2018-09-21_14:09:50"
vignette: >
  %\VignetteIndexEntry{Model Report}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
if (!exists("model")) {
  model <- ""
}
if (!exists("model_dir")) {
  model_dir <- "results/evaluated-scenarios-2018-09-21_14:09:50/baseline-2018-09-21_14:09:50"
}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, 
                      message = TRUE, cache = TRUE,
                      fig.path = file.path("..", model_dir, "plots/"),
                      cache.path = file.path("..", model_dir, "reports", "model_report/"),
                      fig.align = 'center',
                      fig.length = 8, fig.width = 8, eval = TRUE)
```

- Load packages required for analysis

```{r packages, include = FALSE}
library(tidyverse)
library(rbi.helpers)
library(prettypublisher)
library(ModelTBBCGEngland)
```

- Load the fitted model and prior draws.

```{r load-model}
if (class(model) %in% "character") {
  model <- ModelTBBCGEngland::read_libbi(file.path(model_dir, "libbi", "posterior"))
}

priors <- readRDS(file.path(model_dir, "data", "prior-params.rds"))
model
```

- Get the rejection rates of parameters

```{r}
traces <- coda::mcmc(rbi::get_traces(model))

coda::rejectionRate(traces)
```

- Evaluate Posterior Traces

```{r posterior-traces, fig.width = 16, fig.height = 16}
plot_param(model, scales = "free", plot_type = "trace", burn_in = 0) + theme(legend.position = "none")
```

- Plot overview of prior and posterior densities

```{r prior-posterior-overview, fig.height = 16, fig.width = 16}
plot_param(model, prior_params = priors, scales = "free")
```

- Table of priors and posteriors 


```{r param-overview, fig.height = 16, fig.width = 16}
param_sum <- plot_param(model, prior_params = priors, plot_data = FALSE) %>% 
  group_by(Distribution, parameter, length) %>% 
  summarise(median = median(value), 
            lll = quantile(value, 0.025), 
            hhh = quantile(value, 0.975)) %>% 
  mutate(value = pretty_ci(median, lll, hhh, sep = ", ")) %>%
  group_by(Distribution, parameter) %>% 
  mutate(Parameter = case_when(max(length) > 1 ~ paste(parameter, 1:n(), sep = "-"),
                               TRUE ~ as.character(parameter))
         ) %>% 
  ungroup %>% 
  dplyr::select(Distribution, Parameter, value) %>% 
  spread(key = "Distribution", value = "value") %>% 
  dplyr::select(Parameter, Prior, Posterior)

saveRDS(param_sum, file.path(model_dir, "data", "sum_prior_posterior.rds"))
knitr::kable(param_sum)
```

- Calculate incidence sensitivity to parameters at the maximum time fitted to.

```{r}
inc_sens <- test_sensitivity(model, obs = "YearlyInc")

saveRDS(inc_sens, file.path(model_dir, "data", "incidence_sensitivity.rds"))
knitr::kable(inc_sens)
```

- Plot incidence sensitivity to parameters

```{r incidence-sensitivity, fig.height = 16, fig.width = 8}
plot_sensitivity(inc_sens) 
```

- Calculate childhood incidence sensitivity to parameters at the maximum time fitted to.

```{r}
child_inc_sens <- test_sensitivity(model, obs = "YearlyChildInc", time = NULL)

saveRDS(child_inc_sens, file.path(model_dir, "data", "child_incidence_sensitivity.rds"))
knitr::kable(child_inc_sens)
```

- Plot incidence sensitivity to parameters

```{r childhood-incidence-sensitivity, fig.height = 16, fig.width = 8}
plot_sensitivity(inc_sens) 
```


- Plot an overview of states 

```{r overview-states, fig.height = 16, fig.width = 16}
plot_state(model, summarise = TRUE)
```

- Plot an overview of states zoomed into 1990 onwards

```{r overview-states-1990, fig.height = 16, fig.width = 16}
plot_state(model, summarise = TRUE, start_time = 59)
```

- Plot fitted observed states.

```{r overview-states-1980-obs}
plot_state(model, states = c("YearlyHistPInc", "YearlyInc"), summarise = TRUE, start_time = 49) -> p2

p2
```

- Table fitted states

```{r}
pred_obs <- plot_state(model, states = c("YearlyHistPInc", "YearlyInc", "YearlyAgeInc"), summarise = FALSE, start_time = 59, plot_data = FALSE)

obs <- ModelTBBCGEngland::setup_model_obs(years_of_age = 2000:2015, age_groups = 0:11, con_age_groups = c("children", "adults", "older adults")) %>% 
  bind_rows(.id = "state") %>% 
  mutate(time = time + 1931)

pred_obs <- pred_obs %>% 
  dplyr::filter(Average == "median") %>% 
  mutate(pred_value = pretty_ci(Count, lll, hhh, digits = 0)) %>% 
  left_join(obs, by = c("state" = "state", "time" = "time", "age" = "age")) %>% 
  select(Observation = state, Age = age, Year = time,  `Observed Incidence` = value, `Predicted Incidence` = pred_value) %>% 
  mutate(Year = Year)


sum_obs_table <- pred_obs %>% 
  dplyr::filter(Observation %in% c("YearlyInc", "YearlyHistPInc")) %>% 
  drop_na(`Observed Incidence`) %>% 
  select(-Age) %>% 
  mutate(Observation = case_when(Observation == "YearlyInc" ~ "All TB cases",
                                 Observation == "YearlyHistPInc" ~ "Pulmonary TB cases"))


saveRDS(sum_obs_table, file.path(model_dir, "data", "sum_obs_table.rds"))

age_cases_table <- pred_obs %>% 
  dplyr::filter(Observation %in% c("YearlyAgeInc")) %>% 
  drop_na(`Observed Incidence`) %>% 
  select(-Observation) %>% 
  group_by(Year) %>% 
  mutate(Age = c(paste0(seq(0, 45, 5), "-", seq(4, 49, 5)), "50-69", "70-89")) %>% 
  ungroup


saveRDS(age_cases_table, file.path(model_dir, "data", "age_cases_table.rds"))
```

```{r}
knitr::kable(sum_obs_table)
```

```{r}
knitr::kable(age_cases_table)
```

- Plot concentrated age distributed cases


```{r overview-states-1990-con-age}
plot_state(model, states = c( "YearlyChildInc", "YearlyAdultInc", "YearlyOlderAdultInc"), start_time = 59) + theme(legend.position = "none")
```

- Plot age distributed cases.

```{r overview-states-1990-age}
plot_state(model, states = c("YearlyAgeInc"), start_time = 59) + theme(legend.position = "none")
```
